---
title: "scRNASeq_Liver"
author: "André Barros & Helena Cabaço"
date: last-modified
execute:
    echo: true
    warning: false
editor: visual
format:
  html:
    toc: true
    standalone: true
    embed-resources: true
    code-fold: true
    code-tools: true
---

```{r setup, include=F}
knitr::opts_chunk$set(root.dir = "C:/Users/asbarros/OneDrive - FUNDAÇÃO GIMM - GULBENKIAN INSTITUTE FOR MOLECULAR MEDICINE/ADA/ZZ_Tutorials/AutumSchool_scRNASeq/Group_Project/")

library(Seurat)
library(tidyverse)
library(sctransform)
library(crayon)

library(igraph) # for networks
library(pheatmap) # for simple heatmaps

library(presto) # fast DE
library(gprofiler2) # gene set enrichment

library(liana) # cell-cell communication

```

# scRNASeq Analysis Liver - Steady State vs NAFLD

This notebook is focused on the analysis of a scRNASeq dataset of liver, comparing samples from controls (steady state) and NAFLD.

```{r import}
setwd("C:/Users/asbarros/OneDrive - FUNDAÇÃO GIMM - GULBENKIAN INSTITUTE FOR MOLECULAR MEDICINE/ADA/ZZ_Tutorials/AutumSchool_scRNASeq/Group_Project/")

sce = readRDS("00_Dataset/MouseLiverNAFLD.RDS")
sce

orig_metadata = read.table("01_OriginalMetadata/GSE166504_cell_metadata.20220204.tsv",sep="\t",header=T) %>%
  mutate(UniqueID = paste0(FileName,"_",CellID))

sce@meta.data = sce@meta.data %>%
              mutate(UniqueID = paste0(FileName,"_",CellID))

orig_metadata = orig_metadata %>% 
  filter(UniqueID %in% sce@meta.data$UniqueID)


#all(orig_metadata$UniqueID == sce@meta.data$UniqueID)

sce@meta.data$celltype = orig_metadata$CellType


metadata = sce@meta.data

table(metadata$animal,metadata$sort,metadata$time)
table(metadata$capture,metadata$sort,metadata$time)

cat(paste0(bold("\nTotal Number of Genes with >0 counts: "), sum(rowSums(sce@assays$RNA) > 0)," (", 
           round(100*sum(rowSums(sce@assays$RNA) > 0) / nrow(sce@assays$RNA),2),"% of full dataset)")) 

```

The dataset contains \~31k cells, with \~25k features (genes) - out of those, only ~18k (72%) of the genes have counts.

We have hepatocytes & non-parenchimal cells (NPCs), distributed into three groups: Chow (negative control), 15 weeks post-treatment and 30 weeks post-treatment. We have replicates (2-3) per timepoint, per cell type. The capture of the samples was also performed in two separated captures (Capture1 and Capture2).

**Note:** What is the age of the 'Chow' group? In reality, there are changes that happen to the liver throughout age; therefore, the control here is not optimal.

To simplify our work, we are going to keep only the samples *from 30 weeks and from Chow (Negative Control)*:

```{r}

sce@meta.data$Filt = ifelse(sce@meta.data$time=="15weeks",F,T)

sce_30vCh = subset(sce, subset=Filt)

sce = sce_30vCh

sce_30vCh
#################

metadata = sce@meta.data

table(metadata$animal,metadata$sort,metadata$time)
table(metadata$capture,metadata$sort,metadata$time)

cat(paste0(bold("\nTotal Number of Genes with >0 counts: "), sum(rowSums(sce@assays$RNA) > 0)," (", 
           round(100*sum(rowSums(sce@assays$RNA) > 0) / nrow(sce@assays$RNA),2),"% of full dataset)")) 
```

We are now working with ~19K cells, now with ~17.5k features (~70%) with genes with >0 counts. 

## Quality Control

```{r QC1, fig.height=6,fig.width=8}
# Establish the default assay
DefaultAssay(sce) <- 'RNA'

# Add column for the percentage of reads associated with mitochondrial genes - Not found in our dataset

#sce = PercentageFeatureSet(sce, col.name = "percent.mt", assay = "RNA", pattern = "^mt-") 

# Check the plot for features, counts & percentage of mitochondrial genes
VlnPlot(sce, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2,log=T)
VlnPlot(sce, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2,log=T,pt.size = 0)

```

We could not find any mitochondrial genes in the dataset (based on the `^mt-` classification). The expectation is that these genes were filtered out before the raw data was deposited online.

These violin plots are more informative in log scale (especially for the lower cutoffs). Looking at the `nCount_RNA`, we can also observe that there was already some pre-filtering performed in this dataset. For this, we also observe less sequencing depth in NPC's compared with hepatocytes, which is a good indication that cut-offs should be performed per sorted cell type.

We have two batch variables, 'Animal' and 'Capture'. Considering the pre-processing should be done on a *per sample* level (if not in multiplex), let us check if these variables have any influence in the filtering, we should perform violin plots according with these variables:

```{r QC2, fig.height=10,fig.width=8}
f1 = VlnPlot(sce, features = c("nFeature_RNA"), log=T,group.by = "sort",split.by = "animal",pt.size = 0.5)+scale_fill_hue()+
  theme_bw()+
  theme(plot.title=element_text(hjust = 0.5))

f2 = VlnPlot(sce, features = c("nCount_RNA"), log=T,group.by = "sort",split.by = "animal",pt.size=0.5)+scale_fill_hue()+
  theme_bw()+
  theme(plot.title=element_text(hjust = 0.5))

f1 + f2

#####################################################################################################################

f3 = VlnPlot(sce, features = c("nFeature_RNA"), log=T,group.by = "sort",split.by = "capture")+scale_fill_hue()+
  theme_bw()+
  theme(plot.title=element_text(hjust = 0.5))

f4 = VlnPlot(sce, features = c("nCount_RNA"), log=T,group.by = "sort",split.by = "capture")+scale_fill_hue()+
  theme_bw()+
  theme(plot.title=element_text(hjust = 0.5))

f3 + f4

```

For the 'Capture' variable, a customized approach *per capture* is not necessary; on the other hand, there seems to exist a pattern per 'Animal' in our dataset (in fact, we can observe already some performed filtering). Specifically for `nFeature_RNA` in hepatocytes, we can perform a more hard filtering:

-   For Animal 1, hepatocytes can be filtered at 650 features;
-   For Animal 2, the filtering is good,
-   For Animal 3, hepatocytes can be filtered at 650 features;

```{r QC3 & filt, fig.height=6,fig.width=8}

metadata= metadata %>%
  mutate(QC_filt = case_when(
    sort=="NPC" ~ TRUE,
    sort=="Hepatocyte" & animal=="Animal1" & nFeature_RNA>650 ~ TRUE,
    sort=="Hepatocyte" & animal=="Animal2" ~ TRUE,
    sort=="Hepatocyte" & animal=="Animal3" & nFeature_RNA>650 ~ TRUE,
    TRUE ~ FALSE
  ))

sce$QC_filt1 = metadata$QC_filt

VlnPlot(sce, features = c("nFeature_RNA"), log=T,group.by = "sort",split.by = "QC_filt1")+scale_fill_hue()+
  theme_bw()+
  theme(plot.title=element_text(hjust = 0.5))


sce_qcfilt = subset(sce, subset=QC_filt1)



cat(paste0(bold("\nNumber of cells kept: "), ncol(sce_qcfilt@assays$RNA)," (", 
           ncol(sce@assays$RNA) - ncol(sce_qcfilt@assays$RNA)," cells excluded)"))
```

In total, we only discarded additional **63** cells - a small number like this was expected, considered the dataset had been previously filtered.

### Normalization

In the normalization step, we are going to perform `sctransform`. As stated in the paper, it is possible to use this function to account for batch effects - in our case, 'animal' and 'capture'; however, this approach is not fully consensusal - There is some discussing on how useful is this (https://github.com/satijalab/seurat/issues/7866) and if we should do this normalization step *per sample*.

#### `sctransform` on full dataset

According with `sctransform` paper, we have to perform two rounds of normalization - first, to account for library size, and only afterwards the batch effect.

```{r norm}
cat("\nRunning the first normalization step - with just per-sample metrics\n")
sce_norm1 = SCTransform(sce_qcfilt, verbose = F,
                        vars.to.regress = c("nFeature_RNA", "nCount_RNA"),  
                        variable.features.n = 5000,
                        seed.use = 123)

DefaultAssay(sce_norm1) = "SCT"

cat("\nRunning the second normalization step - with per-sample metrics and batch effect\n")
sce_norm2 = SCTransform(sce_norm1, verbose = F,
                        vars.to.regress = c("nFeature_RNA", "nCount_RNA","animal","capture"),  
                        variable.features.n = 5000,
                        seed.use = 123)

```

## Dimensionality Reduction

### Principal Component Analysis (PCA)

First step is to determine the number of principal components (PCs) that should be included for UMAPs

```{r PCA1}

sce_pca = RunPCA(sce_norm2, verbose = T, assay = "SCT", npcs = 50)

#DimPlot(sce_pca, reduction = "pca", dims = 1:2)

ElbowPlot(sce_pca, ndims = 50)

```
Probably, we are safe to use only 30 PC's for the UMAP.

```{r,fig.width=12,fig.height=6}

PCA_df = sce_pca@reductions$pca@cell.embeddings %>% 
  as.data.frame() %>% mutate(UniqueID = rownames(.))

res_PCA = data.frame(PC=NA,sort=NA,cond=NA,r2=NA)

for (i in 1:30){
  
  tb = data.frame(sce_pca@meta.data, PC = PCA_df[,i]) %>%
    mutate(batch=paste0(animal,"_",capture))
  
  tb_list = split(tb,tb$sort)
  
  m1_hep = lm(PC ~ animal,data=tb_list$Hepatocyte)
  m2_hep = lm(PC ~ capture,data=tb_list$Hepatocyte)
  m3_hep = lm(PC ~ batch,data=tb_list$Hepatocyte)
  m4_hep = lm(PC ~ celltype,data=tb_list$Hepatocyte)
  
  m1_NPC = lm(PC ~ animal,data=tb_list$NPC)
  m2_NPC = lm(PC ~ capture,data=tb_list$NPC)
  m3_NPC = lm(PC ~ batch,data=tb_list$NPC)
  m4_NPC = lm(PC ~ celltype,data=tb_list$NPC)

  temp = data.frame(PC=i,
                    sort=rep(c("Hepatocytes","NPC"),each=4),
                    cond=rep(c("Animal","Capture","Batch","CellType"),times=2),
                    r2 = c(performance::r2(m1_hep)$R2_adjusted,performance::r2(m2_hep)$R2_adjusted,performance::r2(m3_hep)$R2_adjusted,
                           performance::r2(m4_hep)$R2_adjusted,
                           performance::r2(m1_NPC)$R2_adjusted,performance::r2(m2_NPC)$R2_adjusted,performance::r2(m3_NPC)$R2_adjusted,
                           performance::r2(m4_NPC)$R2_adjusted))
  res_PCA = rbind(res_PCA,temp)
  
}

res_PCA %>%
  filter(!is.na(PC)) %>%
  ggplot(.,aes(x=PC,y=r2,color=cond))+
    geom_point(size=2,alpha=.7)+
    geom_line(aes(group=cond),linewidth=1)+
    facet_wrap(~sort,nrow=1)+
    theme_bw()+
    ylim(min(res_PCA$PC),1)+
    theme(axis.title.x=element_blank(),axis.text = element_text(size=12),
        plot.title = element_text(hjust=0.5),strip.background=element_rect(fill="black"),
        strip.text = element_text(color="white",size=16),
        legend.position = "bottom",text = element_text(size=16),legend.title=element_blank(),axis.text.y=element_text(size=14))
```
Looking at the results:
- The biggest R2 values are definitely for the CellType, which makes sense. Interestingly, this variable seems to be more relevant for NPCs;
- There are some effects in early PC's with some relevant R2 for animal (in hepatocytes) and for Capture (in NPCs)

This could be an indications we should perform some kind of batch correction (via integration).

### UMAP

```{r,fig.width=12,fig.height=6}

sce_umap <- RunUMAP(sce_pca, dims = 1:30, verbose = F)

sce_umap@meta.data = 
    sce_umap@meta.data %>%
        mutate(batch=paste0(animal,"_",capture))


DimPlot(sce_umap, reduction = "umap",group.by = "animal",split.by = "time")
DimPlot(sce_umap, reduction = "umap",group.by = "capture",split.by = "time")
DimPlot(sce_umap, reduction = "umap",group.by = "batch",split.by = "time")
DimPlot(sce_umap, reduction = "umap",group.by = "celltype",split.by = "time")

```

The UMAPs produced showed us:
- Cell types seem pretty well defined, as expected;
- There are some rotations & clusters that seem associated with capture (mostly for hepatocytes and endothelial cells). This is an indication that we should perform some integration & batch correction in our dataset, taking 'capture' into consideration.

For the sake of time, we are going to filter our dataset only for 'Capture1'.

## Filtering for 'Capture1'

```{r}

sce_filt = subset(sce_qcfilt,subset=capture=="Capture1")

sce = sce_filt

sce_filt
#################

metadata = sce@meta.data

table(metadata$animal,metadata$sort,metadata$time)
table(metadata$capture,metadata$sort,metadata$time)

cat(paste0(bold("\nTotal Number of Genes with >0 counts: "), sum(rowSums(sce@assays$RNA) > 0)," (", 
           round(100*sum(rowSums(sce@assays$RNA) > 0) / nrow(sce@assays$RNA),2),"% of full dataset)")) 

```
We get only ~11k samples now, with ~17K genes with >0 counts (~68% of the dataset).

```{r}

sce_norm = SCTransform(sce, verbose = F,
                        vars.to.regress = c("nFeature_RNA", "nCount_RNA"),  
                        variable.features.n = 5000,
                        seed.use = 123)

saveRDS(sce_norm,"Capt_norm_sce.RDS")

rm(list=ls())

sce = readRDS("Capt_norm_sce.RDS")

```

## Dimensionality Reduction - Only Capture1

### Principal Component Analysis (PCA)

First step is to determine the number of principal components (PCs) that should be included for UMAPs

```{r norm_sep_2}

sce_pca = RunPCA(sce, verbose = T, assay = "SCT", npcs = 50)

#DimPlot(sce_pca, reduction = "pca", dims = 1:2)

ElbowPlot(sce_pca, ndims = 50)
```
We are keeping the number of PC's at 40 - to be a bit more permissive.

### UMAP

```{r,fig.width=12,fig.height=6}

sce_umap <- RunUMAP(sce_pca, dims = 1:40, verbose = F)

sce_umap@meta.data = 
    sce_umap@meta.data %>%
        mutate(batch=paste0(animal,"_",capture))


DimPlot(sce_umap, reduction = "umap",group.by = "animal",split.by = "time")
DimPlot(sce_umap, reduction = "umap",group.by = "celltype",split.by = "time")

```

Although the integration is not perfect, we consider it is safe to proceed with the analysis.

We are going to check the UMAP, divided by cell type:

```{r, fig.width=16,fig.height=10}
DimPlot(sce_umap, reduction = "umap",group.by = "time",split.by = "celltype",ncol = 4)
```

## Downstream Analysis

### Gene Set Enrichment Analysis

```{r}
table(sce$time,sce$celltype)
```

From the data here, we see that the number of cells is considerably reduced in several conditions. For those which have =<5 in one of the groups, we will not considered it. 


```{r}

Idents(sce) = sce$time

cells_df = table(sce$time,sce$celltype) %>% as.data.frame() %>% group_by(Var2) %>% summarise(keep=all(Freq>5))

comp_l = cells_df$Var2[cells_df$keep==T]

de_cond_l = list()

for(n in comp_l){
  
  cat(paste0("\nTaking care of: ",n," - diff. analysis\n"))
    cells = sce$celltype %in% n
    print(table(sce[,cells]$celltype, sce[,cells]$time))
    
    de_cond_l[[n]] = FindMarkers(sce[,cells], ident.1 = "30weeks", ident.2 = "Chow")
}

```


```{r}
allgenes = rownames(sce) # all genes, to serve as background

res_GSEA = data.frame(comp=NA,term_name=NA,intersection_size=NA,p_value=NA)

for (n in names(de_cond_l)) {
  
    cat(paste0("\nTaking care of: ",n," - GSEA\n"))
  
  mk_clusters = de_cond_l[[n]]

  degenes = mk_clusters %>%
    arrange(-avg_log2FC) 
  
  cat(paste0("\n -> Up-regulated"))

  gprof_up = gost(rownames(degenes), ordered_query = T, 
                custom_bg = allgenes, # always set a custom bg!
                sources = "GO:BP", 
                correction_method = "g_SCS", # more stringent than FDR, 
                organism = "mmusculus", user_threshold = 0.05)
  
  degenes = mk_clusters %>%
    arrange(avg_log2FC) 
  
  cat(paste0("\n -> Down-regulated"))
  
  gprof_down = gost(rownames(degenes), ordered_query = T, 
                custom_bg = allgenes, # always set a custom bg!
                sources = "GO:BP", 
                correction_method = "g_SCS", # more stringent than FDR, 
                organism = "mmusculus", user_threshold = 0.05)
  
  temp = gprof_up$result %>%
    rbind(gprof_down$result %>% mutate(intersection_size=-intersection_size)) %>%
    filter(p_value<0.01) %>%
    top_n(n = 10,wt=-log10(p_value)) %>%
    mutate(comp=n) %>%
    select(comp,term_name,intersection_size,p_value)
    
  
res_GSEA = rbind(res_GSEA,temp)
}

```

```{r,fig.width=15,fig.height=22}
res_GSEA %>% 
  mutate(signal=ifelse(intersection_size<0,"down","up")) %>%
  filter(!is.na(comp)) %>%
  ggplot(aes(x=-log10(p_value),y=term_name,color=signal))+
    geom_segment(aes(x = 0, xend =-log10(p_value), y = term_name,yend=term_name), 
                 color = "black")+
    geom_point(size = 4)+
    scale_colour_manual(values = c("down"="darkblue","up"="darkred"))+
    facet_wrap(~comp,ncol=2,scales="free_y")+
    scale_y_discrete(labels = scales::label_wrap(20)) +
    theme_bw()+
    labs(title = "GO:BP Enrichment Analysis")+
    xlab("-log10(p-value)")+
    ylab("GO Terms (BP)")+
    theme(strip.background=element_rect(fill="black"),
          strip.text = element_text(color="white",size=12),
          plot.title = element_text(size=16,hjust=0.5,face="bold"),
          plot.subtitle = element_text(size=12,hjust=0.5,face="italic"),
          legend.position="bottom")
```

### Cell-Cell Communication

```{r}
srat_n = sce[,sce$time=="Chow"]
srat_t = sce[,sce$time=="30weeks"]

cat(paste0("\nProcessing Chow cells\n"))

cell_tab = table(srat_n$celltype)
cells_keep = colnames(srat_n)[srat_n$celltype %in% names(cell_tab)[cell_tab>=10]]
srat_n = srat_n[,cells_keep]
print(sort(unique(srat_n$celltype)))

cat(paste0("\nProcessing 30weeks cells\n"))

cell_tab = table(srat_t$celltype)
cells_keep = colnames(srat_t)[srat_t$celltype %in% names(cell_tab)[cell_tab>=10]]
srat_t = srat_t[,cells_keep]
print(sort(unique(srat_t$celltype)))

cat(bold(paste0("\nDifferences between cell types in conditions\n")))
cat(paste0("\n-> Only for Chow: ",unique(srat_n$celltype)[!unique(srat_n$celltype) %in% srat_t$celltype]))
cat(paste0("\n-> Only for 30weeks: ", as.character(paste0(unique(srat_t$celltype)[!unique(srat_t$celltype) %in% srat_n$celltype],collapse=", "))))

```

We have cells that are specific for each time-point/condition. To make the comparisons more comparable, let's exclude the cell types not overlapping.

```{r}
overlap = intersect(unique(srat_n$celltype),unique(srat_t$celltype))

cat(paste0("\nProcessing Chow cells\n"))
cells_keep = colnames(srat_n)[srat_n$celltype %in% overlap]
srat_n = srat_n[,cells_keep]
print(sort(unique(srat_n$celltype)))

cat(paste0("\nProcessing 30weeks cells\n"))
cell_tab = table(srat_t$celltype)
cells_keep = colnames(srat_t)[srat_t$celltype %in% overlap]
srat_t = srat_t[,cells_keep]
print(sort(unique(srat_t$celltype)))

```

Now, we will proceed to LIANA

```{r}
liana_n = liana_wrap(srat_n, idents_col = "celltype", 
                     method = "cellphonedb", resource = c("MouseConsensus"),
                     permutation.params = list(nperms=500, parallelize=F, workers=4)) # you may need to adjust these depending on your computer
liana_t = liana_wrap(srat_t, idents_col = "celltype", 
                     method = "cellphonedb", resource = c("MouseConsensus"),
                     permutation.params = list(nperms=500, parallelize=F, workers=4)) # you may need to adjust these depending on your computer
```

**Normal Conditions (Chow)**

```{r,fig.width=8,fig.height=6}
liana_n_filt = liana_n[liana_n$pvalue<=0.01,]
liana_t_filt = liana_t[liana_t$pvalue<=0.01,]

q20lr = quantile(c(liana_n_filt$lr.mean, liana_t_filt$lr.mean), 0.20)
liana_n_filt = liana_n_filt[liana_n_filt$lr.mean>q20lr,]
liana_t_filt = liana_t_filt[liana_t_filt$lr.mean>q20lr,]

liana_n_filt_wrapped <- liana_n_filt %>%
  mutate(
    source = str_wrap(source, width = 20),
    target = str_wrap(target, width = 20)
  )


heat_freq(liana_n_filt_wrapped, grid_text=T, font_size = 10)
```

**30 weeks**

```{r,fig.width=8,fig.height=6}

#################################################################

liana_t_filt_wrapped <- liana_t_filt %>%
  mutate(
    source = str_wrap(source, width = 20),
    target = str_wrap(target, width = 20)
  )


heat_freq(liana_t_filt_wrapped, grid_text=T, font_size = 10)
```


```{r,fig.width=12,fig.height=8}
liana_n_filt %>%
  group_by(source,ligand) %>%
  summarise(Freq=n()) %>%
  top_n(n = 10,Freq) %>%
  ggplot(aes(x=Freq,y=ligand))+
    geom_segment(aes(x = 0, xend =Freq, y = ligand,yend=ligand), 
                 color = "black")+
    geom_point(size = 4,color="darkblue")+
    facet_wrap(~source,scales="free_y")+
    theme_bw()+
    labs(title = "Top Ligands per Cell Type",subtitle = "Normal Conditions")+
    xlab("Freq.")+
    ylab("Ligands")+
    theme(strip.background=element_rect(fill="black"),
          strip.text = element_text(color="white",size=12),
          plot.title = element_text(size=16,hjust=0.5,face="bold"),
          plot.subtitle = element_text(size=12,hjust=0.5,face="italic"),
          legend.position="bottom")


liana_n_filt %>%
  group_by(target,receptor) %>%
  summarise(Freq=n()) %>%
  top_n(n = 10,Freq) %>%
  ggplot(aes(x=Freq,y=receptor))+
    geom_segment(aes(x = 0, xend =Freq, y = receptor,yend=receptor), 
                 color = "black")+
    geom_point(size = 4,color="darkred")+
    facet_wrap(~target,scales="free_y")+
    theme_bw()+
    labs(title = "Top Receptors per Cell Type",subtitle = "Normal Conditions")+
    xlab("Freq.")+
    ylab("Ligands")+
    theme(strip.background=element_rect(fill="black"),
          strip.text = element_text(color="white",size=12),
          plot.title = element_text(size=16,hjust=0.5,face="bold"),
          plot.subtitle = element_text(size=12,hjust=0.5,face="italic"),
          legend.position="bottom")

```


```{r,fig.width=12,fig.height=8}
liana_t_filt %>%
  group_by(source,ligand) %>%
  summarise(Freq=n()) %>%
  top_n(n = 10,Freq) %>%
  ggplot(aes(x=Freq,y=ligand))+
    geom_segment(aes(x = 0, xend =Freq, y = ligand,yend=ligand), 
                 color = "black")+
    geom_point(size = 4,color="darkblue")+
    facet_wrap(~source,scales="free_y")+
    theme_bw()+
    labs(title = "Top Ligands per Cell Type",subtitle = "30weeks")+
    xlab("Freq.")+
    ylab("Ligands")+
    theme(strip.background=element_rect(fill="black"),
          strip.text = element_text(color="white",size=12),
          plot.title = element_text(size=16,hjust=0.5,face="bold"),
          plot.subtitle = element_text(size=12,hjust=0.5,face="italic"),
          legend.position="bottom")


liana_t_filt %>%
  group_by(target,receptor) %>%
  summarise(Freq=n()) %>%
  top_n(n = 10,Freq) %>%
  ggplot(aes(x=Freq,y=receptor))+
    geom_segment(aes(x = 0, xend =Freq, y = receptor,yend=receptor), 
                 color = "black")+
    geom_point(size = 4,color="darkred")+
    facet_wrap(~target,scales="free_y")+
    theme_bw()+
    labs(title = "Top Receptors per Cell Type",subtitle = "30weeks")+
    xlab("Freq.")+
    ylab("Ligands")+
    theme(strip.background=element_rect(fill="black"),
          strip.text = element_text(color="white",size=12),
          plot.title = element_text(size=16,hjust=0.5,face="bold"),
          plot.subtitle = element_text(size=12,hjust=0.5,face="italic"),
          legend.position="bottom")

```

**Overlap**

```{r,fig.width=12,fig.height=8}

liana_n_filt = liana_n_filt %>% mutate(Comb = paste0(source,"-",ligand," to ", target," - ",receptor))
liana_t_filt = liana_t_filt %>% mutate(Comb = paste0(source,"-",ligand," to ", target," - ",receptor))

cat(paste0("Nr. of common ligand pairs: ",length(liana_n_filt$Comb[liana_n_filt$Comb %in% liana_t_filt$Comb]),
           " (",round(100*length(liana_n_filt$Comb[liana_n_filt$Comb %in% liana_t_filt$Comb])/nrow(liana_n_filt),2),"% of Normal Interactions, ",
           round(100*length(liana_n_filt$Comb[liana_n_filt$Comb %in% liana_t_filt$Comb])/nrow(liana_n_filt),2),"% of 30Weeks Interactions)"))

liana_n_filt %>%
  filter(!Comb %in% liana_t_filt$Comb) %>%
  group_by(source,ligand) %>%
  summarise(Freq=n()) %>%
  top_n(n = 5,Freq) %>%
  ggplot(aes(x=Freq,y=ligand))+
    geom_segment(aes(x = 0, xend =Freq, y = ligand,yend=ligand), 
                 color = "black")+
    geom_point(size = 4,color="darkblue")+
    facet_wrap(~source,scales="free_y")+
    theme_bw()+
    labs(title = "Top Ligands per Cell Type",subtitle = "Exclusive for Normal Conditions")+
    xlab("Freq.")+
    ylab("Ligands")+
    theme(strip.background=element_rect(fill="black"),
          strip.text = element_text(color="white",size=12),
          plot.title = element_text(size=16,hjust=0.5,face="bold"),
          plot.subtitle = element_text(size=12,hjust=0.5,face="italic"),
          legend.position="bottom")


liana_n_filt %>%
    filter(!Comb %in% liana_t_filt$Comb) %>%
  group_by(target,receptor) %>%
  summarise(Freq=n()) %>%
  top_n(n = 5,Freq) %>%
  ggplot(aes(x=Freq,y=receptor))+
    geom_segment(aes(x = 0, xend =Freq, y = receptor,yend=receptor), 
                 color = "black")+
    geom_point(size = 4,color="darkred")+
    facet_wrap(~target,scales="free_y")+
    theme_bw()+
    labs(title = "Top Receptors per Cell Type",subtitle = "Exclusive for Normal Conditions")+
    xlab("Freq.")+
    ylab("Ligands")+
    theme(strip.background=element_rect(fill="black"),
          strip.text = element_text(color="white",size=12),
          plot.title = element_text(size=16,hjust=0.5,face="bold"),
          plot.subtitle = element_text(size=12,hjust=0.5,face="italic"),
          legend.position="bottom")

####################################################################################################

liana_t_filt %>%
  filter(!Comb %in% liana_n_filt$Comb) %>%
  group_by(source,ligand) %>%
  summarise(Freq=n()) %>%
  top_n(n = 5,Freq) %>%
  ggplot(aes(x=Freq,y=ligand))+
    geom_segment(aes(x = 0, xend =Freq, y = ligand,yend=ligand), 
                 color = "black")+
    geom_point(size = 4,color="darkblue")+
    facet_wrap(~source,scales="free_y")+
    theme_bw()+
    labs(title = "Top Ligands per Cell Type",subtitle = "Exclusive for 30weeks")+
    xlab("Freq.")+
    ylab("Ligands")+
    theme(strip.background=element_rect(fill="black"),
          strip.text = element_text(color="white",size=12),
          plot.title = element_text(size=16,hjust=0.5,face="bold"),
          plot.subtitle = element_text(size=12,hjust=0.5,face="italic"),
          legend.position="bottom")


liana_t_filt %>%
    filter(!Comb %in% liana_n_filt$Comb) %>%
  group_by(target,receptor) %>%
  summarise(Freq=n()) %>%
  top_n(n = 5,Freq) %>%
  ggplot(aes(x=Freq,y=receptor))+
    geom_segment(aes(x = 0, xend =Freq, y = receptor,yend=receptor), 
                 color = "black")+
    geom_point(size = 4,color="darkred")+
    facet_wrap(~target,scales="free_y")+
    theme_bw()+
    labs(title = "Top Receptors per Cell Type",subtitle = "Exclusive for 30weeks")+
    xlab("Freq.")+
    ylab("Ligands")+
    theme(strip.background=element_rect(fill="black"),
          strip.text = element_text(color="white",size=12),
          plot.title = element_text(size=16,hjust=0.5,face="bold"),
          plot.subtitle = element_text(size=12,hjust=0.5,face="italic"),
          legend.position="bottom")

```

````{r,fig.width=6,fig.height=6}

liana_n_excl = liana_n_filt %>%
    filter(!Comb %in% liana_t_filt$Comb) %>%
      mutate(source=case_when(
        source=="Kupffer cells" ~ "KC",
        source=="Monocyte/Monocyte derived macrophage" ~ "Mon",
        source=="Endothelial cells" ~ "EC",
        source=="DCs" ~ "DCs",
        source=="B" ~ "B",
        source=="Stellate cells" ~ "SCs",
        source=="Hepatocytes" ~ "Hep",
        source=="T"~ "T",
        source=="Hepatic progenitor cells" ~ "HPCs",
        TRUE ~ NA
      ),target=case_when(
        target=="Kupffer cells" ~ "KC",
        target=="Monocyte/Monocyte derived macrophage" ~ "Mon",
        target=="Endothelial cells" ~ "EC",
        target=="DCs" ~ "DCs",
        target=="B" ~ "B",
        target=="Stellate cells" ~ "SCs",
        target=="Hepatocytes" ~ "Hep",
        target=="T"~ "T",
        target=="Hepatic progenitor cells" ~ "HPCs",
        TRUE ~ NA
      ))
      


par(cex = 0.8)  # Reduce all text to 80% of default
chord_freq(liana_n_excl)
title("Normal Conditions", cex.main = 1.5)


###################################################################################

liana_t_excl = liana_t_filt %>%
    filter(!Comb %in% liana_n_filt$Comb) %>%
      mutate(source=case_when(
        source=="Kupffer cells" ~ "KC",
        source=="Monocyte/Monocyte derived macrophage" ~ "Mon",
        source=="Endothelial cells" ~ "EC",
        source=="DCs" ~ "DCs",
        source=="B" ~ "B",
        source=="Stellate cells" ~ "SCs",
        source=="Hepatocytes" ~ "Hep",
        source=="T"~ "T",
        source=="Hepatic progenitor cells" ~ "HPCs",
        TRUE ~ NA
      ),target=case_when(
        target=="Kupffer cells" ~ "KC",
        target=="Monocyte/Monocyte derived macrophage" ~ "Mon",
        target=="Endothelial cells" ~ "EC",
        target=="DCs" ~ "DCs",
        target=="B" ~ "B",
        target=="Stellate cells" ~ "SCs",
        target=="Hepatocytes" ~ "Hep",
        target=="T"~ "T",
        target=="Hepatic progenitor cells" ~ "HPCs",
        TRUE ~ NA
      ))
      


par(cex = 0.8)  # Reduce all text to 80% of default
chord_freq(liana_t_excl)
title("30weeks", cex.main = 1.5)

````

## An additional look - Olfactory Receptors

```{r, fig.width=10,fig.height=8}

sce <- PercentageFeatureSet(sce, pattern = "^Olfr", col.name = "pct_olfr",assay = "RNA")
counts <- sce@assays$RNA$counts

olfr_genes <- rownames(counts)[grepl("^Olfr", rownames(counts))]
cat(paste0("\nNr. of Olfr Genes: ",length(olfr_genes),"\n"))

VlnPlot(sce, features = c("pct_olfr"), group.by="celltype",split.by="time",sort = T)+
  ggtitle("Percentage of Reads in Olfr Genes")+
  theme(legend.position = "bottom") # these two metrics are also useful in log scale

cat("\nNr. of Olfrs with expression per cell: \n")
table(Matrix::rowSums(counts[olfr_genes, , drop = FALSE] > 0))

sce$nFeature_Olfr <- Matrix::colSums(counts[olfr_genes, , drop = FALSE] > 0)

VlnPlot(sce, features = c("nFeature_Olfr"),group.by="celltype",split.by="time",sort = T)+
  ggtitle("Nr. of Olfr Genes w/ expression per cell")+
  theme(legend.position = "bottom") # these two metrics are also useful in log scale

rowSums(counts[olfr_genes, , drop = FALSE] > 0) [rowSums(counts[olfr_genes, , drop = FALSE] > 0)>0]

```

```{r,fig.width=10,fig.height=8}

Olfr_interest = names(rowSums(counts[olfr_genes, , drop = FALSE] > 0) [rowSums(counts[olfr_genes, , drop = FALSE] > 0)>0])

FeaturePlot(sce_umap, features = Olfr_interest)
```


#SessionInfo
```{r}
sessionInfo()
```